{How do we copy and paste?\\The semantic drift of quotations in blogspace}

### 2015-02-10 brains-introduction.tex
corrections generales jusqu'a la section 3

Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics. A first stream is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution. It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}) following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the counterpart of the former outside of the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: ``A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention. The producer and the user of a mental representation are one and the same person. A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}''.}

One of the core claims of this literature consists in emphasizing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized. Sperber's cultural epidemiology opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
Cultural epidemiology postulates that this conceptual evolution can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and some of the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
Despite some recent modeling attempts (\hbox{e.g.} \citet{Claidiere07}), the development of quantitative measurements relying on the concept of cultural attractors has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analyzed in an extensive manner.

Another research stream deals with rather macroscopic studies of knowledge diffusion. Here, one of the focal points is that not all knowledge gets propagated identically along the same routes, within the same communities, at the same pace. The various approaches usually feature a minimalistic description of cognitive processes, strongly reminiscent of biological epidemiology (a single, atomic piece of information may or may not be adopted by each individual).  
This research program nonetheless exhibits a significant empirical track record --- largely owing to a recent avalanche of observable \emph{in vivo} data which, for a good decade now, have mainly come from online interaction contexts.
While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
Some authors could describe for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.

These latter studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis). They nonetheless show us the added value of using these rapidly growing records %--~given the already significant and rapidly growing importance of our online interactions~--
%and, %more broadly, large textual corpora \cam{google n-grams??} 
towards radically improving the empirical understanding of (individual-level) cultural evolution processes.

@@

We focus on simple linguistic modifications, thereby connecting our research to the broader psycholinguistic literature.
To deal with robust and simple cultural representations, we paid attention to the evolution of quotations.
While these verbatim public representations should in theory not suffer any alterations as they are produced anew (as opposed to more elaborate expressions and opinions, not identified as quoted utterances), empirical observation shows that they are in fact quite often transformed.
We will in particular exhibit a non-trivial process by which individual words in quotations are replaced.
We will uncover some of the semantic and structural characteristics of these words and the substitutions they undergo.
We moreover contend that using this type of data is equivalent to a large-scale psycholinguistic experiment and at the same time constitutes the first step towards building empirically realistic models of cultural evolution.

>>>>

Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics. A first stream is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution. It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}) following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the counterpart of the former outside of the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
%\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: ``A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention. The producer and the user of a mental representation are one and the same person. A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}''.}

One of the core claims of this literature consists in emphasizing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized. Sperber's cultural epidemiology classically opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
Cultural epidemiology postulates that this conceptual evolution can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and some of the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
Despite some recent modeling attempts (\hbox{e.g.} \citet{Claidiere07}), the development of quantitative measurements relying on the concept of cultural attractors has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analyzed in an extensive manner.

Another research stream deals with rather macroscopic studies of knowledge diffusion. Here, one of the focal points is that not all knowledge gets propagated identically along the same routes, within the same communities, at the same pace. The various approaches usually feature a minimalistic description of cognitive processes, strongly reminiscent of biological epidemiology (a single, atomic piece of information may or may not be adopted by each individual).  
This research program nonetheless exhibits a particularly interesting empirical track record --- largely owing to a recent avalanche of observable \emph{in vivo} data which, for a good decade now, have mainly come from online interaction contexts.
While these information trails are not records of ``physical'' inter-individual interactions (in the sense of ``real life'' interactions), they still constitute a wealth of observations on the dynamics of public --~albeit online~-- representations.
Some authors could describe for instance the propagation of cultural artifacts across social networks such as blogspace \citep{Gruhl04}, the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media \citep{Leskovec09}, or the reciprocal influence between the social network topology and the distribution of issues \citep{Cointet09}.

These latter studies are at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis) and are relatively remote from cognitive science; for a significant part, they rely rather marginally on specific social science theories. They nonetheless show us the added value of using these rapidly growing records %--~given the already significant and rapidly growing importance of our online interactions~--
%and, %more broadly, large textual corpora \cam{google n-grams??} 
towards radically improving the empirical understanding of (individual-level) cultural evolution processes.

@@

We focus on simple linguistic modifications, thereby connecting our research to the broader psycholinguistic literature.
To deal with robust and simple cultural representations, we paid attention to the evolution of quotations.
While these verbatim public representations should in theory not suffer any alterations when they are produced anew (as opposed to more elaborate expressions and opinions, not identified as quoted utterances), empirical observation shows that they are occasionally transformed.
We will in particular exhibit a non-trivial process by which individual words in quotations are replaced.
We will uncover some of the semantic and structural characteristics of these words and the substitutions they undergo.
In some way that using this type of data is equivalent to a large-scale psycholinguistic experiment and at the same time constitutes a first step towards building empirically realistic models of cultural evolution.

### 2015-02-10 brains-protocol.tex
corrections generales jusqu'a la section 3

The number of synonyms is related to a notion of the connectivity of words in a semantic network.  To go a bit further in this direction, we appraise the possible role of network-based variables which have received special attention in the recent related literature, following the blooming interest in networks from many disciplines over the last decade.

\medskip
Aside from classical psycholinguistic measures, we thus considered more recently studied variables based on semantic network properties.
We relied on the ``free association'' norms collected by~\citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations. Free association (FA) norms record the words that come to mind when someone is presented with a given cue (that is the ``free association'' task). As \citet{Nelson04} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
%Following \citet{Griffiths07}, we first consider the directed weighted network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
% \rk{we're in fact using the unweighed version of the network. Why?}
%Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where words are nodes and edges are directed from cue to target word whenever a target word is being produced when this particular cue word was presented.
Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where nodes are words and edges are directed from cue to target word whenever the considered target word was produced in response to the considered cue word.
This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyze is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.

@@

\begin{itemize}
    \item \textbf{Degree centrality}, measured by the number of cues for which a given word is triggered as a target, and a corresponding generalized measure, node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by~\citet{Griffiths07}.
    In the present case the two measures are quasi-perfectly correlated, and measure polysemy by taking into account the structure of the entire network.
    %However in the present case there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, which will lead us to favour the latter later on. Word pagerank on the FA network had already been used by~\citet{Griffiths07}; it may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
    \item \textbf{Betweenness centrality}, another measure of node centrality describing the extent to which a node connects otherwise remote areas of the network~\citep{free:set}.
    This quantity tells us if some words behave like unavoidable waypoints on the path associating one word to another.
    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\cite{watt-coll}, computed on the undirected version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual vs. possible edges between a node's neighbors.}
    This tells us if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
\end{itemize}

@@

The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($0.8$).
The analysis showed a better prediction effect of the number of phonemes over the number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.

@@

\begin{figure}[h]
    \centering
    \subfloat[Possible paths from occurrence to occurrence]{
	    \def\svgwidth{\linewidth}
	    \small
	    \input{images/substitution-temporal-binning-a.pdf_tex}
	    \label{fig:substitution-temporal-binning-a}
	}
	\hfill \\
	\subfloat[Binned quotation family with majority rule]{
	    \def\svgwidth{\linewidth}
	    \small
	    \input{images/substitution-q_max.pdf_tex}
	    \label{fig:substitution-q_max}
	}
	\caption{{\bf Temporal binning of quotation families.} In bin~3, $q'$ now holds the majority and is considered the unique basis for the last occurrence of $q$ (in bin~4). This is despite the fact that $q$ also appears in bin~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario). Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence in bin~3.}
    \label{fig:substitution-temporal-binning}
\end{figure}

@@

Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.  In practice, from the 838k initial occurrences spread into 19,621 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 2,873 real substitutions obeying to this model. From these substitutions we remove those featuring stop words, minor spelling changes (e.g. center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers; this eventually yields 476 valid substitutions.

>>>>

The number of synonyms is related to a notion of the word connectivity in a semantic network.  To go a bit further in this direction, we appraise the possible role of network-based variables which have received special attention in the recent related literature, following the blooming interest in networks from many disciplines over the last decade.

\medskip
%, we thus considered more recently studied variables based on semantic network properties.
We relied on the ``free association'' (FA) norms collected by~\citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations. FA norms record the words that come to mind when someone is presented with a given cue% (that is the ``free association'' task)
. As \citet{Nelson04} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
%Following \citet{Griffiths07}, we first consider the directed weighted network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
% \rk{we're in fact using the unweighed version of the network. Why?}
%Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where words are nodes and edges are directed from cue to target word whenever a target word is being produced when this particular cue word was presented.
Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where nodes are words and edges are directed from cue to target word whenever the considered target word was produced in response to the considered cue word.
This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyze is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.

@@

\begin{itemize}
    \item \textbf{Degree centrality}, measured by the number of cues for which a given word is triggered as a target, and a corresponding generalized measure, node \emph{pagerank}~\citep{Page99}, which has already been used on the FA network by~\citet{Griffiths07}.
    In the present case these two polysemy-related measures are quasi-perfectly correlated. %, and measure polysemy by taking into account the structure of the entire network.
    %However in the present case there is a quasi-perfect correlation between node incoming degree and node \emph{pagerank}~\citep{Page99}, which will lead us to favour the latter later on. Word pagerank on the FA network had already been used by~\citet{Griffiths07}; it may be interpreted as a generalized and recursive measure of word polysemy: central nodes in the pagerank sense are words often selected as targets when presented with cues themselves often selected as targets, and so on recursively.
    \item \textbf{Betweenness centrality}, another measure of node centrality describing the extent to which a node connects otherwise remote areas of the network~\citep{free:set}.
    This quantity tells us if some words behave like unavoidable waypoints on association chains connecting one word to another.
    \item \textbf{Clustering coefficient}, which measures the extent to which a node belongs to a local aggregate of tightly connected nodes~\cite{watt-coll}, computed on the undirected version of the FA network.\footnote{The Clustering coefficient is formally defined as the ratio between the number of actual vs. possible edges between a node's neighbors.}
    This tells us if a word belongs more or less to a local aggregate of equivalent words (from a ``free association'' point of view).
\end{itemize}

@@

The number of phonemes and the number of syllables naturally exhibit a strong linear correlation ($0.8$).
The analysis showed a better prediction effect of the number of phonemes over the number of syllables, which is consistent with~\citet{nick-diss}, and we therefore chose to only present results for the former.\marginnote{C: which analysis, ours? (if yes, the notion of ``prediction effect'' wouldn't appear to be very understandable at this stage of the paper)}

@@

\begin{figure}[h]
    \centering
    \subfloat[Possible paths from occurrence to occurrence]{
	    \def\svgwidth{\linewidth}
	    \small
	    \input{images/substitution-temporal-binning-a.pdf_tex}
	    \label{fig:substitution-temporal-binning-a}
	}
	\hfill \\
	\subfloat[Binned quotation family with majority rule]{
	    \def\svgwidth{\linewidth}
	    \small
	    \input{images/substitution-q_max.pdf_tex}
	    \label{fig:substitution-q_max}
	}
	\caption{{\bf Temporal binning of quotation families.} $q$ and $q'$ are two versions of a quotation belonging to the same cluster.  In the bottom panel (b), $q'$ holds the majority in the 3rd bin and is considered the unique basis for the last occurrence of $q$ (in the 4th bin). This is despite the fact that $q$ also appears in bin~3 alongside $q'$, and despite it having appeared earlier at the very beginning of the quotation family (indeed in the situation shown in Figure~\ref{fig:substitution-q_max}, this seems to be the most likely scenario). Conversely, if $q$ had been the most frequent quote in bin~3, the last occurrence of $q$ in bin~4 would have been considered a faithful copy of the occurrence of $q$ in bin~3.}
    \label{fig:substitution-temporal-binning}
\end{figure}

@@

\medskip
Put shortly, such a model defines how many times quote occurrences can be counted as substitutions: in Figure~\ref{fig:substitution-q_max}, occurrences of $q'$ on the 20th are counted as substitutions, whereas the occurrences on the 21st are not.  In practice, from the 838k initial occurrences spread into 19,621 classes of quotes, with significant redundancy (many quotes are indeed simple duplicates), we manage to mine 2,873 real substitutions obeying to this model. From these substitutions we remove those featuring stop words, minor spelling changes (e.g. center/centre, November/Nov, Senator/Sen), abbreviations, spelled out numbers; this eventually yields 476 valid substitutions.
